### 2021-07-21 paper/paper.tex
enable eprint fields
draft of abstract

@@

\begin{abstract}
TODO: abstract (at most 600 characters, written in plain English with no symbol
nor formula + references etc.) %TODO
\end{abstract}

@@

\begin{itemize}
  \item Basic design of Trixi.jl and its capabilities
  \item PID comparison of Fluxo and Trixi.jl
  \item Discussion on Julia for numerical simulation science
  \item Prepare repro repo including scripts for Trixi.jl and Fluxo
  \item Cite \cite{bezanson2017julia,schlottkelakemper2021purely,rackauckas2017differentialequations,revels2016forward,ranocha2021sbp}
  \item Change title to something like ``Adaptive high-order numerical simulations
  of hyperbolic PDEs with Trixi.jl: A case study of Julia for scientific computing''?
  \item Restructure non-conservative stuff and add MHD benchmarks?
\end{itemize}

@@

@@

\section{Design of Trixi.jl and overview of its features}

@@

FLUXO\footnote{\url{https://gitlab.com/project-fluxo/fluxo}} is an open source
MPI Fortran code implementing discontinuous Galerkin methods on unstructured hex
meshes in three space dimensions for advection diffusion equations. It also
implements the same kind of modern DG methods based on flux differencing
to achieve entropy conservation/dissipation or kinetic energy preservation.
Thus, both Trixi.jl and FLUXO share a common set of features, enabling a reasonable
comparison of their performance.

Here, we compare the serial performance of Trixi.jl and FLUXO when solving a
hyperbolic PDE in three space dimensions on curvilinear hex meshes.

@@

The results of this comparison are visualized in Figure~\ref{fig:PID-Euler}.
Clearly, Trixi.jl is more than 2x faster than the Fortran code FLUXO
while both implement the same numerical methods. This demonstrates the
suitability of Julia for this kind of simulation-focused scientific computing.

\begin{figure}
\centering
  \begin{subfigure}{\linewidth}
    \begin{tikzpicture}[
      font=\footnotesize
      ]
      \begin{axis}[
          xlabel={Polynomial degree},
          % ylabel={Time per RHS call and DOF (seconds)},
          ylabel={Time/RHS/DOF {[sec]}},
          width=\textwidth,
          height=0.66\textwidth,
          legend columns=-1,
          % legend to name=place_legend_here,
          legend pos=north west,
          ymin=0.0,
          ymax=2.2e-6,
          grid=major,
        ]
        \addplot[blue, mark=*, solid] % FLUXO, weak form
          table [x index=0, y index=3]{../data/pids_euler_weak.dat};
          \addlegendentry{FLUXO}
%         \addplot[orange, mark=square*, solid] % TreeMesh, weak form
%           table [x index=0, y index=1]{../data/pids_euler_weak.dat};
%           \addlegendentry{\texttt{TreeMesh} (Trixi.jl)}
        \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
          table [x index=0, y index=2]{../data/pids_euler_weak.dat};
          % \addlegendentry{\texttt{StructuredMesh} (Trixi.jl)}
          \addlegendentry{Trixi.jl}
        \node at (axis cs:13,3.5e-7) {Weak form};

@@

Finally, the different mesh types available in Trixi.jl allow further optimizations.
For example, if the computational domain is essentially a cube (or square, line)
without need for curvilinear coordinates, the Cartesian \lstinline{TreeMesh} can
be used. Depending on the relative computational effort of the discretization
(such as the polynomial degree and the choice of the volume terms), the
Cartesian \lstinline{TreeMesh} can be ca.\ \SI{10}{\percent}--\SI{25}{\percent}
more efficient than the curvilinear meshes in Trixi.jl.

>>>>

% Macros
\newcommand{\trixi}{Trixi.jl}

@@

\begin{abstract}
% Abstract: at most 600 characters, written in plain English with no symbol
% nor formula + references etc.
We present Trixi.jl, a Julia library for adaptive high-order numerical simulations
of hyperbolic partial differential equations (PDEs). Utilizing Julia's strengths,
Trixi.jl is extendable, easy to use, and fast. We describe the main design choices
that allow satisfying these goals and compare Trixi.jl with an established open
source Fortran code for hyperbolic PDEs using the same numerical methods.
We conclude with an assessment of Julia for simulation-focused scientific
computing, an area that is still dominated by traditional high-performance
computing languages such as C, C++, and Fortran.
\end{abstract}

@@

\begin{itemize}
  \item Basic design of \trixi and its capabilities
  \item PID comparison of Fluxo and \trixi
  \item Discussion: Julia for numerical simulation science
  \item Prepare repro repo including scripts for \trixi and Fluxo
  \item Cite \cite{bezanson2017julia,schlottkelakemper2021purely,rackauckas2017differentialequations,revels2016forward,ranocha2021sbp}
  \item Change title to something like ``Adaptive high-order numerical simulations
  of hyperbolic PDEs with \trixi: A case study of Julia for scientific computing''?
  \item Restructure non-conservative stuff and add MHD benchmarks?
\end{itemize}

@@

\trixi\footnote{https://github.com/trixi-framework/Trixi.jl} provides adaptive
high-order numerical simulations of hyperbolic partial differential equations (PDEs).
in Julia \cite{bezanson2017julia}.

@@

\section{Design of \trixi and overview of its features}

@@

FLUXO\footnote{\url{https://gitlab.com/project-fluxo/fluxo}} is an open source
MPI Fortran code implementing discontinuous Galerkin methods on unstructured hex
meshes in three space dimensions for advection diffusion equations. It also
implements the same kind of modern DG methods based on flux differencing
to achieve entropy conservation/dissipation or kinetic energy preservation.
Thus, both \trixi and FLUXO share a common set of features, enabling a reasonable
comparison of their performance.

Here, we compare the serial performance of \trixi and FLUXO when solving a
hyperbolic PDE in three space dimensions on curvilinear hex meshes.

@@

The results of this comparison are visualized in Figure~\ref{fig:PID-Euler}.
Clearly, \trixi is more than 2x faster than the Fortran code FLUXO
while both implement the same numerical methods. This demonstrates the
suitability of Julia for this kind of simulation-focused scientific computing.

\begin{figure}
\centering
  \begin{subfigure}{\linewidth}
    \begin{tikzpicture}[
      font=\footnotesize
      ]
      \begin{axis}[
          xlabel={Polynomial degree},
          % ylabel={Time per RHS call and DOF (seconds)},
          ylabel={Time/RHS/DOF {[sec]}},
          width=\textwidth,
          height=0.66\textwidth,
          legend columns=-1,
          % legend to name=place_legend_here,
          legend pos=north west,
          ymin=0.0,
          ymax=2.2e-6,
          grid=major,
        ]
        \addplot[blue, mark=*, solid] % FLUXO, weak form
          table [x index=0, y index=3]{../data/pids_euler_weak.dat};
          \addlegendentry{FLUXO}
%         \addplot[orange, mark=square*, solid] % TreeMesh, weak form
%           table [x index=0, y index=1]{../data/pids_euler_weak.dat};
%           \addlegendentry{\texttt{TreeMesh} (\trixi)}
        \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
          table [x index=0, y index=2]{../data/pids_euler_weak.dat};
          % \addlegendentry{\texttt{StructuredMesh} (\trixi)}
          \addlegendentry{\trixi}
        \node at (axis cs:13,3.5e-7) {Weak form};

@@

Finally, the different mesh types available in \trixi allow further optimizations.
For example, if the computational domain is essentially a cube (or square, line)
without need for curvilinear coordinates, the Cartesian \lstinline{TreeMesh} can
be used. Depending on the relative computational effort of the discretization
(such as the polynomial degree and the choice of the volume terms), the
Cartesian \lstinline{TreeMesh} can be ca.\ \SI{10}{\percent}--\SI{25}{\percent}
more efficient than the curvilinear meshes in \trixi.

TODO: Plot of this comparison with the \lstinline{TreeMesh}?
